---
title: "Homework 2"
author: Laura Gomez
output: github_document
---


```{r setup}
library(tidyverse)
library(readxl)

```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df =
  read_xlsx(
    "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls), 
  sports_balls = as.integer(sports_balls)
  )
```

Read in precipiation data! For 2018 and 2017

```{r}
precip_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation. 

```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018 , precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr.Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheers include month precipitation data. 

## Problem 2

Read the NYC Transit data. 

```{r}
library(tidyverse)
nyc_transit_df = 
  read_csv(
    ("./Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")) %>%
  janitor::clean_names() 
```
Clean the dataset to include line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.Convert the entry variable from a character to a logical variable. 

```{r nyc_clean_transit_df}
nyc_clean_transit_df = 
  select(nyc_transit_df, line:vending, ada, -exit_only) %>% drop_na(entry) %>% drop_na(vending) %>%
  mutate(entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE), vending = recode(vending, 'YES' = TRUE, 'NO' =   FALSE))
```

This dataset contains information information related to each entrance and exit for each subway station in NYC. We cleaned and selected for the data to include `r ls(nyc_clean_transit_df)` and removed repeated objects.We converted the data in the **entry** and **vending** to a logical character. There are a total of `r nrow(nyc_clean_transit_df)` rows and `r ncol(nyc_clean_transit_df)` columns in our final dataset. 

The data is not very tidy because the **Route** columns have been NA values scattered throughout the data. The dataset can be further organized into clusters that help condense the data. 

## Code for to calculate below answers
```{r}
## Number of distinct stations
distinct_nyc_df = distinct(nyc_clean_transit_df,station_name, station_latitude, .keep_all = TRUE)
```
There are `r nrow(distinct_nyc_df)` distinct stations. 
```{r}
## Number of stations that are ADA compliant?
ada_compliant_df = filter(distinct_nyc_df, ada == TRUE )
```
There are `r nrow(ada_compliant_df)` stations that are ADA compliant.
```{r}
## Proportion of station entrances / exits without vending allow entrance?

entrance_df = nrow(filter(nyc_transit_df, vending == "NO", entry == TRUE))

no_entrance_df = nrow(filter(nyc_transit_df, vending == "NO", entry == FALSE))

proportion = (entrance_df) / (no_entrance_df)
```
The proportion of station entrance to exists without vending allow entrance is `r nrow(proportion)`. 
```{r}
##How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

### Reformat data so that route number and route name are distinct variables
Route_df = nyc_clean_transit_df %>% 
  mutate(route8 = as.character(route8), route9 = as.character(route9),
  route10 = as.character(route10), route11 = as.character(route11)) %>% 
  pivot_longer(route1:route11, names_to = "Route_name" , values_to = "Route_number")

Route_distinct_df = distinct(Route_df,Route_name, Route_number, .keep_all = TRUE )

### Number of stations serve the A train.

 #Route_A_distinct_df = nrow(filter(Route_distinct_df, Route_name == "A"))

 #Route_A_distinct_ADAcompliant_df = nrow(filter(Route_A_distinct_df, ada == TRUE))
```

There are r Route_A_distinct_df stations that serve the A train. Within these station r Route_A_distinct_ADAcompliant_df are ADA compliant. 

